Tools for identifying unexpectedly low microfossil count sums

### 2020-06-11 Rmd/count_check_MS.Rmd
add hershop bog

loadd(bird_singleton_summary)
loadd(testate_summ)
loadd(last_chance_summ)
loadd(pollen_summ, pollenA, pollen1234)
loadd(owens_summ1, owens_summ, owens_est_summ)
loadd(diatom1_summ, diatom2_summ, chironomid1_summ, marine1_summ, chironomid2_summ)

@@

The pollen data from the Neotoma database included over `r pollen_summ$n_assemblages` assemblages in `r pollen_summ$n_datasets %>% format()` datasets. 
Nearly all assemblages have singletons (`r pollen_summ$p_singletons`%) and a greatest common divisor of one (`r pollen_summ$p_gcd1`%). 
The assemblages with a greatest common divisor above one are not randomly distributed among the datasets: only `r pollen_summ$dataset_gcd1`% of the datasets have any such assemblages.
Some datasets with a high proportion of assemblages with a greatest common divisor above one are older datasets which may, despite the available metadata, have been digitised with the loss of rare taxa and precision.
For example, one dataset includes `r pollenA$n_samp` assemblages with a median richness of `r pollenA$taxa_median` taxa.
Count sums vary between `r pollenA$count_min` and `r pollenA$count_max`.
The greatest common divisor is three for all assemblages, that is all `r pollenA$n_count` counts in the dataset are divisible by three.
Excluding such datasets would further increase the proportion of assemblages with singletons. 
Some other Neotoma datasets are discussed is a subsequent section.

@@

Some of the pollen datasets in the Neotoma database have an inexplicably high proportion of assemblages with a greatest common divisor greater than one. 
Here I focus on four anonymised datasets produced by one research group. 
All the datasets, which are relatively recent, include information on the spike used to calculate pollen concentration, so cannot have been digitised.
The four datasets have some assemblages with a greatest common divisor of two interspersed, sometimes regularly, amongst assemblages with the expected greatest common divisor of one.
The assemblages with a greatest common divisor of one typically have several singletons (Table \@ref(tab:neotoma-table)).

>>>>

loadd(bird_singleton_summary)
loadd(testate_summ)
loadd(last_chance_summ)
loadd(pollen_summ, hershop_pollen_summ, pollenA, pollen1234)
loadd(owens_summ1, owens_summ, owens_est_summ)
loadd(diatom1_summ, diatom2_summ, chironomid1_summ, marine1_summ, chironomid2_summ)

@@

The pollen data from the Neotoma database included over `r pollen_summ$n_assemblages` assemblages in `r pollen_summ$n_datasets %>% format()` datasets. 
Nearly all assemblages have singletons (`r pollen_summ$p_singletons`%) and a greatest common divisor of one (`r pollen_summ$p_gcd1`%). 
The assemblages with a greatest common divisor above one are not randomly distributed among the datasets: only `r pollen_summ$dataset_gcd1`% of the datasets have any such assemblages.
Some datasets with a high proportion of assemblages with a greatest common divisor above one are older datasets which may, despite the available metadata, have been digitised with the loss of rare taxa and precision.
For example, all but `r hershop_pollen_summ$has_singleton %>% word()` of the `r hershop_pollen_summ$n` assemblages in the archived pollen data from Hershop Bog Texas have a greatest common divisor of two or more (the exception has a single singleton and is otherwise divisible by two). 
Discrepancies between the archived data and the original paper [@Larson_1972] including an incomplete species list and the inclusion of fern spores in the count sum of two hundred grains strongly suggest that these data have been digitised.
For most purposes, such digitised data are perfectly adequate.

Excluding such datasets would further increase the proportion of assemblages with singletons. 
Some other Neotoma datasets are discussed in section \@ref(unexpected-count-data) of this paper.

@@

Some of the pollen datasets in the Neotoma database have an inexplicably high proportion of assemblages with a greatest common divisor greater than one. 
Here I focus on five anonymised datasets.

Dataset pollen1, published includes `r pollenA$n_samp` assemblages with a median richness of `r pollenA$taxa_median` taxa.
Count sums vary between `r pollenA$count_min` and `r pollenA$count_max`.
The greatest common divisor is three for all assemblages, that is all `r pollenA$n_count` counts in the dataset are divisible by three.

All the datasets, which are relatively recent, include information on the spike used to calculate pollen concentration, so cannot have been digitised.
The four datasets have some assemblages with a greatest common divisor of two interspersed, sometimes regularly, amongst assemblages with the expected greatest common divisor of one.
The assemblages with a greatest common divisor of one typically have several singletons (Table \@ref(tab:neotoma-table)).
